### 2020-02-29 doc/icfp20/intro.tex

Artists and theoreticians have attempted to capture and codify these
rule systems, always informally in natural language, and typically
accompanied by examples from the existing literature. The intent is
both to analysize existing music and then to use these principles to
guide the creation of new music, in other words for synthesis.

Starting in the 20th century, computers have become ubiquitous in
music in every area, including sound sythesis, composition and
production (\citep{roads-tutorial}).

[more here]

>>>>

Artists and theoreticians have attempted to capture and codify these
rule systems, informally in natural language, and typically
accompanied by examples from the existing literature. The intent is
both to analyze existing music and then to use these principles to
guide the creation of new music, in other words for synthesis.

Starting in the 20th century, computers have become ubiquitous in
music in every area, including sound sythesis, composition and
production~\citep{roads-tutorial}. In terms of music theory, there has
been a line of recent work on using functional programming for
harmonic
analysis~\citep{magalhaes-harmtrace,dehaas-harmtrace-a,dehaas-harmtrace-b},
harmonization of a melody and generation of melodies based on a
harmonization~\citep{koops-fharm,magalhaes-fcomp} and
counterpoint~\citep{szamozvancev-welltyped}. There is also an
established Haskell library Euterpea~\citep{hudak-haskell} for general music
and sound exploration.

To describe rules of basic harmonic structure,
\citet{magalhaes-harmtrace} and its successors use dependent
types. (TODO: Give an example.) However Haskell currently has limited
support for dependent types, and requires many extentions and tricks
such as the use of singleton types~\citep{eisenberg-singleton} to
simulate a restricted form of dependent types. (TODO: fix wording)

We would like to explore what can be done in the context of music by
using a programming language that offers full dependent types. We use
the language Agda~\citep{norell-phd} since it is fairly mature and
aims to be a functional programming language as well as a proof
assistant. It also features a Haskell FFI so we can take advantage of
existing Haskell libraries, in partiular for MIDI.

Full dependent types allow a fully expressive logic, and it is
tempting to take a standard textbook on music theory such as
\citet{piston-harmony} or \citet{aldwell2018harmony} and attempt to
formalize it in type theory.

-- people are starting to use computers for analysis and synthesis of rule systems
(reference Haskell work). In particular FP, and some rudimentary dep types

-- With full dep types, expressive logic, tempting to formalize all the rules
of some theory book in dep types. Point out difficulties....

-- Music Tools, compare to Euterpea + other Haskell work, mention FARM.

-- In this paper, focus on counterpoint plus basic start to some harmony.

Counterpoint section: contrast old and new methods

Harmony section: Harmonize melody, use counterpoint for voice leading and to contrain harmony.
Mention grid for harmonic analsis, makes use of vectors.

conclusion: easy to express rules, break into parts, reuse in other contexts, extend.
Downsides: hard to do some things (equality, sets, etc), Proof emphasis can get in the way, but
may want to use in the future. Future work: equivalence, ornaments, ML.

TODO: Fix bib style.
